### 2023-04-05 example/DomainDecomposition/EBCellMerge/hoeb_geometry_test/_doc/notes.tex
very ebcm data holders

\begin{itemize}
\item BlobFaceIndex is a collection of EBIndexes
\item BlobVoluIndex is a merged bunch of EBIndex<CELL>
\item Merging and putting into the unstructured format should be
  separate things.
\item Working theory on how the usage pattern for this stuff will work:
  \begin{itemize}
  \item Build EB as normal.
  \item Put EB into irregular description.
    \item while still in irregular description, merge small cells ON
      BOX.
    \item  If we find we cannot merge on box, for now throw an error.
    \item  Later we can figure out Box to Box communication.
  \end{itemize}
\item Notes on memory overhead:
  \begin{itemize}
    \item This could mean  a large increase in the memory load for the
      geometric meta data but this can be managed by doing stuff in a
      carefully ordered manner:
      \begin{itemize}
      \item Build EBIS.
      \item Build BlobSpace using EBIS.  The irregular graph is a
        temporary here that gets deleted.
      \item Delete EBIS.
      \item Now if you build data and operators after this process is
        complete, the meta data overhead should be no worse than EBIS.
      \end{itemize}

  \end{itemize}

\end{itemize}