### 2023-04-18 example/DomainDecomposition/EBCellMerge/hoeb_geometry_test/_doc/paper.tex
lots of refinements on data acquisition for the paper

\documentclass{article}

%\include{macros}

\bibliographystyle{plain}

\usepackage{epsfig}
\usepackage{xcolor}
%%\usepackage{amssymb}
\usepackage{amsmath}

\newcommand{\Abold}{{\bf A}}
\newcommand{\abold}{{\bf a}}
\newcommand{\bbold}{{\bf b}}
\newcommand{\cbold}{{\bf c}}
\newcommand{\dbold}{{\bf d}}
\newcommand{\ebold}{{\bf e}}
\newcommand{\fbold}{{\bf f}}
\newcommand{\Fbold}{{\bf F}}
\newcommand{\gbold}{{\bf g}}
\newcommand{\hbold}{{\bf h}}
\newcommand{\ibold}{{\bf i}}
\newcommand{\jbold}{{\bf j}}
\newcommand{\kbold}{{\bf k}}
\newcommand{\lbold}{{\bf l}}
\newcommand{\mbold}{{\bf m}}
\newcommand{\Mbold}{{\bf M}}
\newcommand{\nbold}{{\bf n}}
\newcommand{\obold}{{\bf o}}
\newcommand{\pbold}{{\bf p}}
\newcommand{\qbold}{{\bf q}}
\newcommand{\rbold}{{\bf r}}
\newcommand{\sbold}{{\bf s}}
\newcommand{\tbold}{{\bf t}}
\newcommand{\ubold}{{\bf u}}
\newcommand{\vbold}{{\bf v}}
\newcommand{\wbold}{{\bf w}}
\newcommand{\xbold}{{\bf x}}
\newcommand{\ybold}{{\bf y}}
\newcommand{\zbold}{{\bf z}}
\newcommand{\bigzbold}{{\bf Z}}
\newcommand{\xspace}{\hspace{2 mm}}

\newcommand{\ahat}{{\hat a}}
\newcommand{\bhat}{{\hat b}}
\newcommand{\chat}{{\hat c}}
\newcommand{\dhat}{{\hat d}}
\newcommand{\ehat}{{\hat e}}
\newcommand{\fhat}{{\hat f}}
\newcommand{\ghat}{{\hat g}}
\newcommand{\hhat}{{\hat h}}
\newcommand{\ihat}{{\hat i}}
\newcommand{\jhat}{{\hat j}}
\newcommand{\khat}{{\hat k}}
\newcommand{\lhat}{{\hat l}}
\newcommand{\mhat}{{\hat m}}
\newcommand{\nhat}{{\hat n}}
\newcommand{\ohat}{{\hat o}}
\newcommand{\phat}{{\hat p}}
\newcommand{\qhat}{{\hat q}}
\newcommand{\rhat}{{\hat r}}
\newcommand{\shat}{{\hat s}}
\newcommand{\that}{{\hat t}}
\newcommand{\uhat}{{\hat u}}
\newcommand{\vhat}{{\hat v}}
\newcommand{\what}{{\hat w}}
\newcommand{\xhat}{{\hat x}}
\newcommand{\yhat}{{\hat y}}
\newcommand{\zhat}{{\hat z}}

\newcommand{\abar}{{\bar {\bf a}}}
\newcommand{\bbar}{{\bar {\bf b}}}
\newcommand{\cbar}{{\bar {\bf c}}}
\newcommand{\dbar}{{\bar {\bf d}}}
\newcommand{\ebar}{{\bar {\bf e}}}
\newcommand{\fbar}{{\bar {\bf f}}}
\newcommand{\gbar}{{\bar {\bf g}}}
%\newcommand{\hbar}{{\bar {\bf h}}}
\newcommand{\ibar}{{\bar {\bf i}}}
\newcommand{\jbar}{{\bar {\bf j}}}
\newcommand{\kbar}{{\bar {\bf k}}}
\newcommand{\lbar}{{\bar {\bf l}}}
\newcommand{\mbar}{{\bar {\bf m}}}
\newcommand{\nbar}{{\bar {\bf n}}}
\newcommand{\obar}{{\bar {\bf o}}}
\newcommand{\pbar}{{\bar {\bf p}}}
\newcommand{\qbar}{{\bar {\bf q}}}
\newcommand{\rbar}{{\bar {\bf r}}}
\newcommand{\sbar}{{\bar {\bf s}}}
\newcommand{\tbar}{{\bar {\bf t}}}
\newcommand{\ubar}{{\bar {\bf u}}}
\newcommand{\vbar}{{\bar {\bf v}}}
\newcommand{\wbar}{{\bar {\bf w}}}
\newcommand{\xbar}{{\bar {\bf x}}}
\newcommand{\ybar}{{\bar {\bf y}}}
\newcommand{\zbar}{{\bar {\bf z}}}

\newcommand{\dx}{{h}}
\newcommand{\nph}{{n + \frac{1}{2}}}
\newcommand{\iph}{{\ibold + \frac{1}{2}\ebold_d}}
\newcommand{\ipmh}{{\ibold \pm \frac{1}{2}\ebold_d}}
\newcommand{\imh}{{\ibold - \frac{1}{2}\ebold_d}}

\newcommand{\half}{\frac{1}{2}}
%\newcommand{\Fbold}{\mathbf{F}}
%\newcommand{\iphed}{{\ibold+\half\ebold^d}}
\newcommand{\deriv}{\partial}
\newcommand{\area}{\mathcal{A}}
\newcommand{\R}[1]{\mathbb{R}^{#1}}
\newcommand{\cf}{{\scriptscriptstyle F}}
\newcommand{\cn}{{\scriptscriptstyle N}}
\newcommand{\sEB}{{\scriptscriptstyle EB}}
\renewcommand{\vec}[1]{\mathbf{#1}} % Vector (bold)
\newcommand{\tens}[1]{\mathbf{#1}} % Tensor (bold)
\newcommand{\mat}[1]{\mathbf{#1}} % Matrix (bold)
\newcommand{\hatvec}[1]{\hat{\vec{#1}}} % Vector with a hat (bold)
\newcommand{\ddt}[1]{\frac{\partial #1}{\partial t}} % partial d/dt
\newcommand{\DDt}[1]{\frac{\text{d}#1}{\text{d}t}} % "total" d/dt
\newcommand{\ddr}[1]{\frac{\partial #1}{\partial r}} % d/dr
\newcommand{\ddx}[1]{\frac{\partial #1}{\partial x}} % d/dx
\newcommand{\ddy}[1]{\frac{\partial #1}{\partial y}} % d/dy
\newcommand{\ddz}[1]{\frac{\partial #1}{\partial y}} % d/dz
\newcommand{\ddxi}[1]{\frac{\partial #1}{\partial x_i}} % d/dx_i
\newcommand{\diverg}[1]{\nabla\cdot#1} % Divergence operator
\newcommand{\curl}[1]{\nabla \times #1} % Curl operator
\newcommand{\dOmega}{\text{d}\Omega} % Volume differential (Omega style)
\newcommand{\dV}{\text{d}V} % Volume differential
\newcommand{\dA}{\text{d}A} % Area differential
\newcommand{\labelEq}[1]{\label{eq:#1}\hbox{\tt #1 \quad}} % Use this to label an equation.
\newcommand{\refEq}[1]{(\ref{eq:#1})}   % Use this to reference an equation.
\newcommand{\labelSec}[1]{\label{sec:#1}} % Use this to label a section.
\newcommand{\refSec}[1]{\S\ref{sec:#1}} % Use this to reference a section.
\newcommand{\labelChap}[1]{\label{chap:#1}} % Use this to label a chapter.
\newcommand{\refChap}[1]{Chapter \ref{chap:#1}} % Use this to reference a chapter.
\newcommand{\labelApp}[1]{\label{app:#1}} % Use this to label an appendix.
\newcommand{\refApp}[1]{Appendix \ref{app:#1}} % Use this to reference an appendix.
\newcommand{\ib}{{\bf{i}}}    % bold italic i
\newcommand{\jb}{{\bf{j}}}    % bold italic j
\newcommand{\kb}{{\bf{k}}}    % bold italic k
\newcommand{\lb}{{\bf{l}}}    % bold italic l
\newcommand{\mb}{{\bf{m}}}    % bold italic m
\newcommand{\ub}{{\bf{u}}}    % bold italic u
\newcommand{\xb}{{\bf{x}}}    % bold italic x
\newcommand{\vb}{{\bf{v}}}    % (bold italic) VoF index
\newcommand{\fb}{{\bf{f}}}    % (bold italic) face index
\newcommand{\pb}{{\bf{p}}}    % (bold italic) multi-index p
\newcommand{\qb}{{\bf{q}}}    % (bold italic) multi-index q
\newcommand{\ebd}{{\bf{e}^d}}
\newcommand{\normal}[1]{\vec{n}_{#1}} % Outward normal.
\newcommand{\order}[1]{\mathcal{O}(#1)} % Order notation

\newcommand{\ed}{{\bf{e}_d}}
\newcommand{\Zbold}{{\bf{0}}}
\newcommand{\unitV}{\mathds{1}}
\newcommand{\eb}{\text{EB}}
\newcommand{\EB}{\text{EB}}
\newcommand{\vol}{\mathcal{V}}
\newcommand{\face}{\mathcal{F}}
\newcommand{\zerobold}{{\bf{0}}}
\newcommand{\xz}{{\bf{x}_0}}

\newcommand{\Dim}{D}
\newcommand{\dif}{\mathrm{d}}
\newcommand{\dt}{{\Delta t}}
\newcommand{\avg}[1]{\overline{#1}}
\newcommand{\avgI}[1]{\overline{#1}_{\ibold}}
\newcommand{\imhed}{{\ibold+\frac{1}{2}\ebold_d}}
\newcommand{\ipmhj}{{\ibold\pm\frac{1}{2}\ebold_d}}
\newcommand{\iphj}{{\ibold+\frac{1}{2}\ebold_d}}
\newcommand{\imhj}{{\ibold-\frac{1}{2}\ebold_d}}

\newcommand{\ipfed}[1]{\ibold+\frac{#1}{2}\ebold_d}
\newcommand{\imfed}[1]{\ibold-\frac{#1}{2}\ebold_d}
\newcommand{\ipmfed}[1]{\ibold\pm\frac{#1}{2}\ebold_d}
\newcommand{\eq}[1]{(\ref{#1})}

\newcommand{\nref}{n_{\text{ref}}}
\newcommand{\phio}{\phi^{\circ}}

\begin{document}

\title{Polynomial expansions of finite volume data in a cut cell context}
\author{N. Overton-Katz \footnotemark[1]
   \and H. Johansen     \footnotemark[1]
   \and D. T. Graves    \footnotemark[1]
   \and D. Devendran    \footnotemark[2]
   \and O. Anteparra    \footnotemark[1]}

\maketitle

\begin{abstract}

  We discuss the mechanics of forming stable local polynomial
  expansions of finite volume data in the context of cut cell grids.
  We show that significant solvability issues can arise from certain weighting
  functions.    A very rudimentary but stable cell merger algorithm is
  presented as we present evidence that these solvability issues may be
  an artifact of small-cell instabilties.

\end{abstract}

\section{Introduction}

Embedded boundary (EB) grids are formed when one passes an  surface
through a Cartesian mesh.    For sufficiently complex geomtries, these methods
are very attactive because grid generation is a solved problem
and not computationally overwhelming even in a moving-boundary context
\cite{MillerTrebotich2012}.

In the current context, we form the cutting surface as the zero
surface of a function  of space $I(\xbold)$, $\xbold \in R^D$.
For smooth ($I$),  moments can be generated to any
accuracy \cite{Schwartz2015}.

EB grids can contain volumes that are arbitrarily small.    This
"small-cell problem'' produces
algorithmic difficulties that generations of researchers have labored
to overcome with great success.   EB methods are used for high speed
compressible flows \cite{Colella2006, Graves2013} and for
slower, incompressible applications \cite{Trebotich2015} and conjugate
heat transfer \cite{Crockett2010}.

Formally, the underlying description of space
is given by rectangular control volumes on a Cartesian mesh
$\Upsilon_\ibold = [(\ibold-\half {\ubold})h, (\ibold+\half
{\ubold})h], \ibold \in \bigzbold^D$, where $D$ is the dimensionality
of the problem, $h$ is the mesh spacing, and ${\ubold}$ is the vector
whose entries are all one (note we use bold font $\ubold = (u_1, \dots, u_d,
\dots, u_D)$ to indicate a vector quantity).
Given an irregular domain $\Omega$, we
obtain control volumes $V_\ibold = \Upsilon_\ibold \bigcap \Omega$ and
faces $A_{\ibold,d\pm} = A_{\ibold \pm \half \ebold_d}$ which are the
intersection of the boundary of $\partial V_\ibold$ with the
coordinate planes $\{{\xbold}:x_d = (i_d \pm \half)h \}$ ($\ebold_d$ is
the unit vector in the $d$ direction).  We also
define $A_{B,\ibold}$ to be the intersection of the boundary of the
irregular domain with the Cartesian control volume: $A_{B,\ibold}
= \partial \Omega \bigcap \Upsilon_\ibold$.

\section {Finite volume notation}

Throughout this paper, we use the following compact notation:
\begin{align*}
(\xbold - \xbar)^\pbold &= \prod\limits^D_{d=1} (x_d - {\bar x}_d)^{p_d} \\
\pbold! &= \prod\limits^D_{d = 1} p_d!
\end{align*}
Given a point in space $\xbar$, and a $D$-dimensional integer vector
$\pbold$, we define $m_v^\pbold(\xbar)$ to be the $\pbold^{th}$
moment of the volume $V$ relative to the point $\xbar$.
\begin{equation}
m_v^\pbold(\xbar)  =  \int\limits_{V} (\xbold - \xbar)^\pbold dV
\label{eqn::volmoment}
\end{equation}

Given a sufficiently smooth function $\psi$, we can approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P_T$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P_T} C^p (\xbar -
\xbar)^p \label{eqn::taylor}
\label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac{1}{p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}  
In finite volume  methods, we define grid data to be
averages over volumes.   So if we discretize the smooth function
$\psi$, the average over the volume $\ibold$ is given by
\begin{equation*}
 V_\ibold <\psi>_\ibold = \int\limits_{V_\ibold} \psi(\xbold) dV
\end{equation*}
If we insert the Taylor expansion \ref{eqn::taylor}, we get a discrete
approximation to the smooth function $\psi$ that is accurate to order $P_T$.
\begin{equation*}
 V_\ibold <\psi>_\ibold = \sum\limits_{p < P_T} C^p m^p. 
\end{equation*}
This forms a local polynomial expansion of $\psi$ around the volume
$\ibold$ expressed in terms moments ($m$) which are the natural
products of grid generation.  Modern, higher order embedded boundary
methods use this description directly to generate stencils
\cite{Overton2022a, Devendran2017, Schwartz2015}.  At the
core, all these algorithms have a similar procedure for finding the
Taylor coefficients $C^p$ which is worth discussing.

The discerning reader will notice that
the equation set used to create $C$ only includes the volume
equations.  Devendran et al. \cite{Devendran2017} also include boundary
condition equations in the equation set.  For simple enough boundary
conditions, this works very well.  There are many equation sets,
however, which do not lend themselves easily to this approach.
Notably, hyperbolic equations often have boundary conditions which are
fomulated in terms of characteristic variables.  Using this
information in the system of equations for Taylor coefficients is
still a matter of research.  We therefore believe that restricting
ourselves to the volume equations can provide insight valuable to
developers of this class of algorithms.

\section{Computing Taylor coefficients and Matrix Conditioning}

Consider a finite volume grid where we know all geometric moments to
order $P_T$.  Given values of $<\psi>$ everwhere on the grid, we wish to
approximate the Taylor coefficients at a volume $\ibold$ given that we
know $\cn_\ibold = \{\jbold\}$, the set of $N_v$ closest neighbors of
the volume $\ibold$.  Using the Taylor expansion at each point gives
us $N_v$ equations for $C^p$
\begin{equation}
  V_\jbold <\psi>_\jbold = \sum\limits_{p < P_T} C^p_\ibold m^p_\jbold.
\end{equation}
Let's say there are $N_p$ Taylor coefficients.  We can force $N_v >
N_p$ by getting a large enough neighborhood.  This means that we have
an overdetermined system of the form
\begin{equation*}
M C = P 
\end{equation*}
where $M = \{\{m^{pi}_\jbold\}\}$, $C=\{C^p\}$,  and $P= \{<\psi_\jbold>\}$.
Since the system is overdetermined, we can introduce a 
meaningful weighting matrix $W$ \footnotemark[3].  Choice of $W$
is crucial to the stability of these algorithms
\footnotemark[4].
We multiply and solve
\begin{equation*}
WMC = WP.
\end{equation*}
Since these matricies are (deliberately) not square, we approximate
the solution using a Moore-Penrose pseudo-inverse:
\begin{equation}
  C = ((WM)^T(WM))^{-1} (WM)^T P.
\label{eqn::getC}  
\end{equation}
This formulation can woven into the computation for stencil
components;  Devendran \cite{Dendran2017}, for example, explains the
process for Poisson's equation.

Unfortunately, the matrix $A \equiv (WM)^T(WM)$ can be poorly conditioned.
That these algorithms can invert a poorly-conditioned matrix to compute
stencil coefficients introduces an interesting type of uncertainty: error in
the very coefficients of a stencil operation.   If the matrix is poorly
conditioned, the coefficients of the stencil can only be known to
limited precision.   This lack of precision is measurable as are its
consequences.

\section{Condition number measurements}

To gain insight into this process, we evaluate the matrix $A \equiv
(WM)^T(WM)$ at every volume in a computational grid and analyze its
eigenvalues.   Given a matrix $A_i$ with the set of eigenvalues
$\{ \lambda_{i}\}$, the inverse of the condition number $I_i$ is given by
\begin{equation*}
  I_i = \frac{\lambda_i|_{min}}{\lambda_i|_{max}},
\end{equation*}
which goes to zero as the maximum and minimum eigenvalues diverge and
the matrix becomes more poorly conditioned.   These eigenvalues are
obtained through  SVD decompostion \footnotemark[7].

First we  measure   an uncut, Cartesian mesh to
provide a baseline without the complications involved with cut cells.
Next we cut the mesh with a very smooth implicit function and do the
same measurements.   Finally we take this cut mesh and merge its small
cells to neighbors to eliminate all cells smaller than a Cartesian
volume to isolate the effects of small cells.

All volumes $V_i$ start with a index $i \in Z^D$.
We start from volume $V_0$.   For 
weighting function, we need to measure $D_{w, V_0}(V_1)$,
the Manhattan disance to the volume $V_1$:
\begin{equation}
  D_{W, V_0}}(V_1) = \sum\limits_{d \in {1...D}} |i_1^d - i_0^d|.
\label{eqn::uptownFunk}
\end{equation}
We define $W_{v_0}(V_1)$, the weighting function used by volume $V_0$ for the
equation describing  volume $V_1$ we use is given by
\begin{equation}
  W_{v_0}(V_1)  \equiv \frac{1}{D^{P_w}_{V_0}(V_1)},
\label{eqn::weightFunk}  
\end{equation}
where $ 0 < P_w < 6$, the weighting power.   We present results for
polynomial order $0 < P_T < 5$.

\subsection{All Regular geometry}

We begin with a motivating example a geometry wherein all cells are uncut.

\subsection{Cut Cell Geometry}
\subsection{Merged Geometry}

\section{Conclusions}
\section{Appendix:Cell Merger Algorithm}

We use a very simple cell merger algorithm.  Given a cut cell
$\ibold$, we define the box containing only that cell
$B_c = B(\ibold, \ibold)$ 
We form a $2^D$ cell rectangular region
by coarsening and then refining $B_c$ to form $B_v = R(C(B_c))$
\footnotemark[6].
After that process, any left over small cells are merged with their
biggest neighbor.    See figures \ref{fig::unmerged2dmap} and
\ref{fig::merged2dmap} for a simple example of how this algorithm
creates fairly chunky grids   \footnotemark[5].

We define the volume fraction $\kappa_\ibold \equiv  V_\ibold/h^D$.
For unmerged grids, $ 0 < \kappa <= 1$.   When merged using this
algorithm, $\kappa$ can be larger than $2^D$.

\begin{figure}
\centerline{\epsfig{figure=_fig/unmerged.ps,width=0.4\linewidth }} 
\caption{Map for EB grid produced by a cutting cirle. $\kappa <= 1$.
  Each cut cell has a unique positive integer.}
\end{figure}

\begin{figure}
\centerline{\epsfig{figure=_figmerged.ps,width=0.4\linewidth } }
\label{fig::merged2dmap}
\caption{Map of EB grid shown in \ref{fig::unmerged2dmap} after using
  cell merger algorithm.
  Each cut cell has a unique positive integer. $\kappa <= 4.1433$}
\end{figure}

\footnotetext[1]{Lawrence Berkeley National Laboratory, Berkeley,
  CA. Research at LBNL was supported financially by the Office of
  Advanced Scientific Computing Research of the US Department of
  Energy under contract number DE-AC02-05CH11231.}
\footnotetext[2]{Ford Motor Company, Sunnyvale, CA.}
\footnotetext[3]{If the matrix $M$ is square, a weighting matrix can
  have no effect and the Moore-Penrose psueudoinverse becomes
  $M^{-1}$.}

\footnotetext[4]{Defining $\xbar_\ibold = 0$, the
  equations for volumes $\jbold$ where $\xbar_\jbold$ target receive
  higher weights.  Diagonal weighting matricies are common.  Usually,
  these algorithms define a distance metric $D(\ibold, \jbold)$ for
  two volumes $\ibold$ and $\jbold$.  Typically they make
  $W_{\jbold,\jbold}$ decrease strongly with increasing 
  $D(\ibold,\jbold)$ to assign higher importance to the equations for volumes
  closer to $\ibold$.  Devendran for example \cite{Devendran2017},
  uses a weighting function that varies with $W_{\ibold,\jbold}
  \approx 1./D(\ibold,\jbold)^5$ }.

\footnotetext[5] {Merging two cells is simply a matter of removing the
  intervening faces, shifting all moments them to a common location
  and adding them together.  No accuracy is lost in this process
  outside of the fact that we have fewer degress of freedom because
  semantically, we are just adding volume integrals over distinct
  volumes, which is exact.}

\footnotetext[6] {We mean a Box as defined in Chombo, a subset of
  $Z^D$.  Coarsening and refining has the standard Chombo semantic
  \cite{ChomboDesign, ChomboDesignEB}.}

\footnotetext[7] {All matrix computations here are performed using
  Eigen \cite{eigenweb}, a beautifully designed and implemented
  package for on-processor linear algebra that is freely available.}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}\
\bibliography{references}

\end{document}