### 2023-05-07 example/DomainDecomposition/EBCM/roadmap.tex
document early and often

\documentclass{article}

\begin{document}

\title{EBCM Directory Roadmap}
\author{
    D. T. Graves    \footnotemark[1]
        }
\date{May 7, 2023 ce.}
\maketitle

\section{hoeb geometry test}

This test makes a single-box EBCM::MetaDataLevel.  It then makes a ton
of data on matrix conditioning in an hoeb/ebcm context.    The
associated technical report is quite complete.

\section{hoeb spmd}

This tests the communications infrastructure for {\tt
Chombo4/src/EBCM}, the bits of Chombo that handle calculations
within the context merged and
unmerged cut cell geometries.   The associated document is complete
but minimal.

\section{hoeb petsc}

This test investigates the eigenvalues and eigenvectors of stencils to
solve Poisson's equation.

\end{document}