### 2023-05-19 example/EBCM/hoeb_geometry/_saved_data/_sphere/notes.tex
moved cell merged examples into their own space

2d:

unmerged output:
nx = 32
max_grid = 32
Cell merging is turned OFF.
making grids
sphere implicit function with radius = 0.45, center = (0.5,0.5), and inside = 1
defining geometry in EB land
makeTheBlobs: number of small cells = 0, min kappa = 0.00880994
maximum (non covered) volume fraction for grids[0] = 1
minimum (non covered) volume fraction for grids[0] = 0.00880994
maximum volume fraction overall = 1
minimum volume fraction overall = 0.00880994
printing time table

merged output:

spencer(101)<hoeb_geometry_test> ./main.exe _inputs/sphere.inputs 
nx = 32
max_grid = 32
Cell merging is turned ON.
making grids
sphere implicit function with radius = 0.45, center = (0.5,0.5), and inside = 1
defining geometry in EB land
makeTheBlobs: number of small cells = 20, min kappa = 0.0661023
maximum (non covered) volume fraction for grids[0] = 4.1433
minimum (non covered) volume fraction for grids[0] = 1
maximum volume fraction overall = 4.1433
minimum volume fraction overall = 1
printing time table