### 2023-05-19 example/EBCM/roadmap.tex

\documentclass{article}

\begin{document}

\title{EBCM Directory Roadmap}
\author{ D. T. Graves }
\date{May 7, 2023 c.e.}
\maketitle

\section{hoeb geometry test}

This test makes a single-box EBCM::MetaDataLevel.  It then makes a ton
of data on matrix conditioning in an hoeb/ebcm context.    The
associated technical report is quite complete.

\section{hoeb spmd}

This tests the communications infrastructure for {\tt
Chombo4/src/EBCM}, the bits of Chombo that handle calculations
within the context of  cut cell geometries where merger is an option.
The associated document is complete but minimal.

\section{hoeb petsc}

This test investigates the eigenvalues and eigenvectors of stencils to
solve Poisson's equation.

\end{document}