### 2023-07-07 example/EBCM/roadmap.tex
updated directory roadmap to reflect current traffic conditions

\documentclass{article}

\begin{document}

\title{EBCM Directory Roadmap}
\author{ D. T. Graves }
\date{July 7, 2023 c.e.}
\maketitle

\section{hoeb geometry test}

This test makes a single-box EBCM::MetaDataLevel.  It then makes a ton
of data on matrix conditioning in an hoeb/ebcm context.    The
associated document (Lawrence Berkeley National Laboratory
Technical Report LBNL-2001528)
technical report is quite complete.

\section{hoeb spmd}

This tests the communications infrastructure for {\tt
Chombo4/src/EBCM}, the bits of Chombo that handle calculations
within the context of  cut cell geometries where merger is an option.
The associated document is complete but minimal.

\section{hoeb petsc}

This test investigates the solvability of the system matrix when
boundary condition equations are included in the system.
The associated document
(Lawrence Berkeley National Laboratory
Technical Report LBNL-2001537)
is complete to the point of being downright
chatty. Does not use petsc so the test name needs work.

\section{hoeb truncation}

This continues the story by investigating the truncation error of
the family of finite volume operators.   This one does use quite a bit
of petsc and includes yet another user interface to the petsc
infrastructure.   The document vaporware as of July 7 2023.

\end{document}